MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-08-07 darmstadt_gscl_2013/article_twitter_normalization.tex
rule-based section for Twitter-specific phenomena added

So, Kukich (\cite{kukich}), for instance, divided all NTN techniques
into six classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

Though both divisions seem to be justified to some extent, it is
difficult to determine using them whether an NTN approach that detects
and restores incorrectly spelled words on the basis of phonetical
\textit{n}-gram statistics should fall into the \textit{n}-gram based,
probabilistic, spell checking or speech recognition class.

@@

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}), Han and Baldwin (\cite{han})
etc. With regard to the type of information inference, practically all
of these methods were supervised with the exceptions of Cook and
Stevenson (\cite{cook}) who claimed to use an unsupervised technique,
and Sproat et al. (\cite{sproat}) who described both a supervised and
an unsupervised method.

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English data. A few exceptions from that are
approaches suggested by Beaufort et al. (\cite{beaufort}) for French
and Oliva et al. (\cite{oliva}) for Spanish. To find out which
peculiarities of ill-formed words are characteristic for German, we
will perform a quantitive and qualitative analysis of unknown words in
German Twitter in the next Section in order to see what kind of NTN
techniques would be more suitable for handling ill-formed words there.

@@

In order to estimate the percentage of unknown words in Twitter
messages, we randomly selected 10,000 tweets from a previously
collected corpus, split them into sentences and tokenized using social
media-aware tokenizer by Christopher Potts
\footnote{\url{http://sentiment.christopherpotts.net/code-data/happyfuntokenizing.py}}.
After skipping all words which did not contain any alphabetic
characters or consisted only of a single letter, we obtained a list of
129,146 tokens. As reference systems for dictionary lookup we used
open source spell checking program \texttt{hunspell}\footnote{Ispell
  Version 3.2.06 (Hunspell Version 1.3.2); dictionary de\_DE.} and
publicly available part-of-speech tagger
\texttt{TreeTagger}\footnote{Version 3.2 with German parameter file
  UTF-8.} (Schmid, \cite{schmid}).

@@

Here once again the question of classification arose. This time with
regard to the reasons, why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to us to be appropriate:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or because they did not exist at the time
    when dictionaries were created. Another reason for inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are either relatively frequent
omissions of characters made by users or, even more often, automatic
truncations of too long messages which are performed by Twitter
service itself.\footnote{As is generally known, Twitter imposes a
  strong restriction on the length of posted messages which can be no
  longer than 140 characters. Upon exceeding this length, tweets get
  automatically truncated to the maximal allowed length.}

@@

A striking outlier of 16.22~\% for slang tokens in column 1 is
explained by the fact that the word ``RT'' which occurred 1,235 times
in our data and was by far the most frequent OOV in analyzed texts,
was recognized as out-of-vocabulary by \texttt{hunspell} but was not
deemed as such by \texttt{TreeTagger}. Luckily, such singleton cases
were rather rare exceptions during the analyses and did not affect
much the remaining classes of OOVs, so that distributions of unknown
tokens for both tools were more or less similar or at least comparable
to a certain extent.

But an even more important conclusion which can be drawn from the
analyzed data is the fact that for both \texttt{TreeTagger} and
\texttt{hunspell}, Twitter-specific phenomena like special tokens,
colloquial expressions or sloppily typed words accounted for more than
a half of all unknown words found in data. Since different classes of
these Twitter phenomena were formed by different processes and also
have different degrees of ambiguity, we will suggest different
normalization procedures for each of them in the next Section.

@@

\subsection{Replacement of Twitter-Specific Phenomena}
Though hashtags are almost ubiquitous in today's tweets and one would
probably think that they are as old as Twitter itself, an admirable
truth would be that hashtags actually existed long before
Twitter-service was introduced. It were Internet Relay Chats (IRC) --
the forebears of all modern social networks -- which used the hash
sign to mark channel names. And it was Chris Messina, a renowned
advocate of open-source community, who borrowed this idea and brought
it to Twitter. The ``hash godfather'', as Chris Messina describes
himself, sent out a Twitter message suggesting that the pound sign
coul be used for organizing groups on Twitter in August 2007. But a
noticeable dissemination of this ``novelty'' among the Twitter
community only happened a few years later, in 2010. ``I would say 2010
was really the year of the hashtag.'' claims Ginger Wilcox, a founder
of the Social Media Marketing Institute (cf. Parker, \cite{parker}).

Luckily for us, both hashtags and @-names which are another
omnipresent phenomenon in Twitter have very clear formal criteria by
which these tokens could be identified and captured. The existence of
such formal cues suggests that these phenomena could best be handled
by means of rule-based methods. A most appropriate of such methods
seemed to be an analog of finite-state transducers (FST).

>>>>

Kukich (\cite{kukich}), for instance, divided all NTN techniques into
six classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

Though both of these divisions seem to be justified to some extent, it
is difficult to determine using them whether an NTN approach that
detects and restores incorrectly spelled words on the basis of
phonetical \textit{n}-gram statistics should fall into the
\textit{n}-gram based, probabilistic, spell checking or speech
recognition class.

@@

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}), Han and Baldwin (\cite{han})
etc. With regard to the type of information inference, most of these
methods were supervised with the exceptions of Cook and Stevenson
(\cite{cook}) and Han and Baldwin (\cite{han}) who claimed to use an
unsupervised technique. Sproat et al. (\cite{sproat}) described in
their article both a supervised and unsupervised approach.

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English data. A few exceptions from that are
approaches suggested by Beaufort et al. (\cite{beaufort}) for French,
and Oliva et al. (\cite{oliva}) for Spanish. To find out which
peculiarities of ill-formed words are characteristic for German, we
will perform a quantitive and qualitative analysis of unknown words in
German Twitter in the next Section in order to see what kind of NTN
techniques would be most suitable for handling such words there.

@@

In order to estimate the percentage of unknown words in Twitter
messages, we randomly selected 10,000 tweets from a previously
collected corpus, split them into sentences and tokenized using social
media-aware tokenizer by Christopher Potts
\footnote{\url{http://sentiment.christopherpotts.net/code-data/happyfuntokenizing.py}}.
After skipping all words which did not contain any alphabetic
characters or consisted only of a single letter, we obtained a list of
129,146 tokens. As reference systems for dictionary lookup we used
open-source spell checking program \texttt{hunspell}\footnote{Ispell
  Version 3.2.06 (Hunspell Version 1.3.2); dictionary de\_DE.} and
publicly available part-of-speech tagger
\texttt{TreeTagger}\footnote{Version 3.2 with German parameter file
  UTF-8.} (Schmid, \cite{schmid}).

@@

Here once again the question of classification arose. This time with
regard to the reasons why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to us to be appropriate to us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or because they did not exist at the time
    when dictionaries were created. Another reason for inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrary to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are either relatively frequent
omissions of characters made by users or even more often automatic
truncations of too long messages which are performed by Twitter
service itself.\footnote{As is generally known, Twitter imposes a
  strong restriction on the length of posted messages which can be no
  longer than 140 characters. Upon exceeding this length, tweets get
  automatically truncated to the maximal allowed length.}

@@

A striking outlier of 16.22~\% for slang tokens in column 1 of the
Table is explained by the fact that the word ``RT'' which occurred
1,235 times in our texts and was by far the most frequent OOV in
analyzed data, was recognized as out-of-vocabulary by
\texttt{hunspell} but was not deemed as such by
\texttt{TreeTagger}. Luckily, such singleton cases were rather rare
exceptions during the analyses and did not affect much the remaining
classes of OOVs, so that distributions of unknown tokens for both
tools were more or less similar or at least comparable to a certain
extent.

But an even more important conclusion which can be drawn from the
analyzed data is the fact that for both \texttt{TreeTagger} and
\texttt{hunspell}, Twitter-specific phenomena like special tokens,
colloquial expressions or sloppily typed words accounted for more than
a half of all unknown words found during the analysis. Since different
classes of these Twitter phenomena were formed by different processes
and also have different degrees of ambiguity, we will suggest
different normalization procedures for each of them in the next
Section.

@@

\subsection{Replacement of Twitter-Specific Phenomena}
In August 2007, Chris Messina, a renowned advocate of open-source
community, sent out a Twitter message suggesting that the pound sign
could be used for organizing groups on Twitter (cf. Parker,
\cite{parker}). The ``hash godfather'', as Chris Messina described
himself, allegedly borrowed this idea from Internet Relay Chats (IRC)
-- the forebears of modern social networks -- which have been using
the ``\#{}''-sign to mark channel names since the early 1990s.

Luckily for us, this ``novelty'' by Messina as well as @-tokens which
are in Twitter to mark names of user account both have strict formal
features by which these tokens could be identified. Other types of
words which are usually considered as OOVs and have unambiguous formal
traits are hyperlinks, e-mail addresses, and smileys. The presence of
such formal criteria suggests that these classes could best be handled
by rule-methods and namely finite-state techniques like finite-state
transducers (FST) (cf. Jurafsky and Martin, \cite{jurafsky},
pp.57-60).

For our purposes, we developed a prototypic Python system analogous to
FSTs in which a set of regular expressions was associated with
corresponding actions performed on matched subgroups. In this system,
absolutely unambiguous tokens like web links and e-mail addresses were
replaced with special words ``\%Mail'' and ``\%Link''
correspondingly. Emoticons were replaced with tokens ``\%PosSmiley'',
``\%NegSmiley'', depending on the type of emotion presumably conveyed
by these expressions. In cases when an emoticon was ambiguous with
regard to its polarity, it was replaced with the artificial word
``\%Smiley''.

Additionally to that, problematic UTF-8 characters like special quotes
and ellipsis which caused incorrect work of tokenizer were converted
to their corresponding ISO-8859-1 equivalents.

For hashtags, we simply stripped off the pound character from the
beginning of such words thus leaving only the alphabetic part of
tokens. A more complicated case turned out to be the @-tokens. For
them, we had to decide depending on the context whether these tokens
played an important syntactic role in sentence or not. In cases they
did, we additionally looked whether the alphabetic part following the
``@'' character was present in dictionary. If it was, we only stripped
the ``at'' character off, otherwise whole word was replaced with token
``\%UserName''. In cases when @-tokens served as addresses or retweet
marks, we deleted them from message since they were of little interest
and importance for further syntactic and semantic analysis and could
hardly be assigned a proper part-of-speech tag or syntactic role.

All the artificially introduced words were added to custom user
dictionaries of \texttt{TreeTagger} and \texttt{hunspell}. For
tagging, we assigned NE tags to all artificial words except for smiley
replacements which got the tag INJ. Additionally, positions and
lengths of all replacements we remembered along with original words
which were replaced and a restoration step to original forms was
performed for certain classes of tokens after tagging.
